---
title: "Comparing model types - synthetic data"
author: "Murray Logan"
date: today
date-format: "DD/MM/YYYY"
format:
  html:
    ## Format
    theme: [default, resources/style.scss]
    css: resources/style.css
    html-math-method: mathjax
    ## Table of contents
    toc: true
    toc-float: true
    ## Numbering
    number-sections: true
    number-depth: 3
    ## Layout
    page-layout: full
    fig-caption-location: "bottom"
    fig-align: "center"
    fig-width: 4
    fig-height: 4
    fig-dpi: 72
    tbl-cap-location: top
    ## Code
    code-fold: false
    code-tools: true
    code-summary: "Show the code"
    code-line-numbers: true
    code-block-border-left: "#ccc"
    code-copy: true
    highlight-style: atom-one
    ## Execution
    execute:
      echo: true
      #cache: true
    ## Rendering
    embed-resources: true
crossref:
  fig-title: '**Figure**'
  fig-labels: arabic
  tbl-title: '**Table**'
  tbl-labels: arabic
engine: knitr
## execute: 
##   cache: true
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
monofont: DejaVu Sans Mono
classoption: a4paper
bibliography: resources/references.bib
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
    cache.lazy = FALSE,
    tidy = "styler"
)
## allow indented chunks
assignInNamespace(".sep.label",
  "^\\ *(#|--)+\\s*(@knitr|----+)(.*?)-*\\s*$",
    ns = "knitr"
)
knitr::read_chunk("../R/synthetic_data.R")
```
## Purpose

This page documents the production of numerous synthetic data sets
that are to be used to compare the performances of various models that
are candidates for GCRMN. Each synthetic dataset is intended to
represent a unique modelling challenge. The following table summarises
the statistical challenge and the path of the synthetic dataset
designed to represent this issue.

| Statistical challenge                             | File path                                |
|---------------------------------------------------|------------------------------------------|
| Replacing poor sites with good sites              | `synthetic/benthos_fixed_locs_obs_1.rds` |
| Replacing poor sites with good sites (small data) | `synthetic/benthos_fixed_locs_obs_2.rds` |
| Gaps in a temporal series of some sites           | `synthetic/benthos_fixed_locs_obs_2.rds` |
| Gaps in a temporal series of all sites            | `synthetic/benthos_fixed_locs_obs_2.rds` |
| Use of covariates                                 | `synthetic/benthos_fixed_locs_obs_2.rds` |

## Preparations

The bulk of the heavy lifting is performed by the `synthos` package.
This package can be installed by issuing the following:

```{r}
#| label: install synthos
#| results: markup
#| eval: false
#| echo: true
#| cache: false
remotes::install_github("open-AIMS/synthos")
```

Load the necessary R libraries

```{r}
#| label: synthetic libraries
#| output: false
#| eval: true
#| warning: false
#| message: false
#| cache: false
```

Establish the global environment and paths

```{r}
#| label: synthetic global parameters
#| results: hide
#| eval: true
#| echo: true
#| cache: false
```
